---
title: "Index"
author: "Erin Morales"
date: '2022-08-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As described above, you'll be creating six different files. Because the purpose of this assignment is to practice collaboration, each section should be completed by a different person (and each person must work on at least one file on their own). We'll be checking the commit history to ensure that each section was pushed by a different account. Here is additional information on each section:

index.Rmd File
In the Markdown document you create, you should run the other scripts to generate the necessary content your report:

Write an introductory paragraph explaining the domain/data set (feel free to borrow from the project proposal)

Write a summary paragraph that includes information calculated by your summary information function

Render your table and describe pertinent observations found in it

Describe the purpose of each chart (e.g., "this chart was intended to show the relationship between..."), render each chart, and describe pertinent observations found in each chart

Summary Information Script
The first file you should save in your project should store summary information in a list. For example:

```{r}

world_happiness_21 <- read.csv('2021.csv')

```


```{r}
library(ggplot2)
library(dplyr)

world_happiness_21 <- read.csv('2021.csv')

ggplot(world_happiness_21, aes(x = Region, y = Economy..logged.GDP.per.Capita., fill = Region)) +
  geom_boxplot()+
  labs(title=" GDP per capita by world region, 2021", x= "Country", y = "Economy (GDP/Capita)") +
  theme(legend.position = "top")+
  coord_flip()

```
 
```{r}
model <- lm(Health..Life.Expectancy. ~ Economy..logged.GDP.per.Capita., data = happiness_df)
summary(model)

# 1/ add text with geom_text, use nudge to nudge the text
ggplot(happiness_df, aes(x=Economy..logged.GDP.per.Capita., y=Health..Life.Expectancy.)) +
  geom_point() + # Show dots
  geom_text(
    label= data$Country,
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T) + stat_smooth(method = "lm", se = FALSE)

hist(model$residuals)
 
```
