---
title: "Index"
author: "Erin Morales"
date: '2022-08-04'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Introduction:

We believe happiness and health go hand in hand. The range of personal, social, economic, and environmental factors that influence health status are known as determinants of health but can also account for a population's happiness. We believe an intersectional lens is vital in identifying key factors that play into an individual's happiness. In turn, using multidimensional data could be crucial in implementing evidence-based policies to serve the public through an intersectional lens better. 

This is specifically a set of data on World Happiness from 2015 to 2022, so it contains a total of 8 csv files within the set. The joy scores and rankings utilize information from the Gallup World Survey. The scores are based on answers to the most life evaluation address inquired within the survey. This file contains the Happiness Score for 153 countries and the factors used to explain the score.


###Summary Paragraph:

In 2021, Afghanistan had the minimum happiness while Finland had the happiness score. Singapore has the highest life expectancy in 2019 and 2020 but also the largest distrust in the government in 2021. Also, in 2021, Croatians had the highest trust in government. The average rate of happiness peaked in 2021 being 5.51, while the highest median was in 2020,  which could indicate more outliers. 
The lowest average happiness was in 2017, being 5.38. The max GDPs tend to be in western Europe, small, high GDP distribution of GDPS, or the Middle East, which has a significant variation in GDP per capita.  Mean health and life expectancy steadily increased yearly, with the healthiest regions primarily being East Asia and Western Europe. The countries with the least freedom were majorly in Central and Eastern Europe, the Middle East, and Sub-Saharan Africa. In contrast, the countries with the most freedom were Western Europe and Central Asia. The number of countries per year varied, peaking in 2015 at 158 countries and the least in 2019 and 2021 with 149 countries. 

###Tables:

```{r}

avg <- function(df) {
  avg_of_year = mean(df$Happiness.Score)
}
average_2015 = avg(wh_2015_df)
average_2016 = avg(wh_2016_df)
average_2017 = avg(wh_2017_df)
average_2018 = avg(wh_2018_df)
average_2019 = avg(wh_2019_df)
average_2020 = avg(wh_2020_df)
average_2021 = avg(wh_2021_df)
# Group 'Happiness_by_Region_2021' into a table
Happiness_by_Region_2021 <- wh_2021_df %>%
  select(Country, Region, Happiness.Score) %>%
  group_by(Region) %>%
  summarize(avg_by_region = mean(Happiness.Score)) %>%
  arrange(desc(avg_by_region))



# Create a function to take the mean of all numeric columns
numeric_mean <- function(df) {
  numeric_mean_df <- df %>%
    summarise_at(c('Happiness.Score', 'Family', 
                   'Health..Life.Expectancy.' , 'Freedom', 
                 'Trust..Government.Corruption.', 'Generosity'),
               mean, na.rm = TRUE)
  return(numeric_mean_df)
}
# Create a table of all the means of numerical columns
all_means_2015 <- numeric_mean(wh_2015_df)
all_means_2016 <- numeric_mean(wh_2016_df)
all_means_2017 <- numeric_mean(wh_2017_df)
all_means_2018 <- numeric_mean(wh_2018_df)
all_means_2019 <- numeric_mean(wh_2019_df)
all_means_2020 <- numeric_mean(wh_2020_df)
all_means_2021 <- numeric_mean(wh_2021_df)

all_means_df <- rbind(all_means_2015, all_means_2016, all_means_2017,
                      all_means_2018, all_means_2019, all_means_2020, 
                      all_means_2021)
all_means_df$year <- c(2015, 2016, 2017, 2018, 2019, 2020, 2021)
all_means_df <- select(all_means_df, 7, 1, 2, 3, 4, 5, 6)



# Create a function to take the max of all numeric columns
numeric_mins <- function(df) {
  numeric_mins_df <- df %>%
    summarise_at(c('Happiness.Score', 'Family', 
                   'Health..Life.Expectancy.' , 'Freedom', 
                   'Trust..Government.Corruption.', 'Generosity'),
                 min, na.rm = TRUE)
  return(numeric_mins_df)
}

min_2015 <- numeric_mins(wh_2015_df)
min_2016 <- numeric_mins(wh_2016_df)
min_2017 <- numeric_mins(wh_2017_df)
min_2018 <- numeric_mins(wh_2018_df)
min_2019 <- numeric_mins(wh_2019_df)
min_2020 <- numeric_mins(wh_2020_df)
min_2021 <- numeric_mins(wh_2021_df)

all_mins_df <- rbind(min_2015, min_2016, min_2017, min_2018, 
                     min_2019, min_2020, min_2021)
all_mins_df$year <- c(2015, 2016, 2017, 2018, 2019, 2020, 2021)
all_mins_df <- select(all_mins_df, 7, 1, 2, 3, 4, 5, 6)


numeric_max <- function(df) {
  numeric_max_df <- df %>%
    summarise_at(c('Happiness.Score', 'Family', 
                   'Health..Life.Expectancy.' , 'Freedom', 
                   'Trust..Government.Corruption.', 'Generosity'),
                 max, na.rm = TRUE)
  return(numeric_max_df)
}

max_2015 <- numeric_max(wh_2015_df)
max_2016 <- numeric_max(wh_2016_df)
max_2017 <- numeric_max(wh_2017_df)
max_2018 <- numeric_max(wh_2018_df)
max_2019 <- numeric_max(wh_2019_df)
max_2020 <- numeric_max(wh_2020_df)
max_2021 <- numeric_max(wh_2021_df)

all_maxs_df <- rbind(max_2015, max_2016, max_2017, max_2018, 
                     max_2019, max_2020, max_2021)
all_maxs_df$year <- c(2015, 2016, 2017, 2018, 2019, 2020, 2021)
all_maxs_df <- select(all_maxs_df, 7, 1, 2, 3, 4, 5, 6)


# View all tables
View(Happiness_by_Region_2021)
View(all_means_df)
View(all_mins_df)
View(all_maxs_df)

```

The aggregated tables here shown are supposed to represent the combined data from 2015-2021, a total of 7 years. Then The first table is shown to have the top 10 world happiness averages throughout those 7 years. The mean, min, and max table shows the top 7 countries in their respective category. This given data shows the harsh contrast of given data set and how factors such as freedom and generosity may contribute to country's happiness and overall well being.    



###Chart summary:

```{r}

ggplot(world_happiness_21, aes(x = Region, y = Economy..logged.GDP.per.Capita., fill = Region)) +
  geom_boxplot()+
  labs(title=" GDP per capita by world region, 2021", x= "Country", y = "Economy (GDP/Capita)") +
  theme(legend.position = "top")+
  coord_flip()

```
In this data visualization there are box plots representing the GDP capita by each world region. This visualization is helpful in identifying the different interquartile ranges and allows us to easily locate the mins and maxs of each region to compare globally.
 
 
```{r}
GDPvsHealthExpectancy <- lm(Health..Life.Expectancy. ~ Economy..logged.GDP.per.Capita., data = happiness_df)
summary(model)

ggplot(happiness_df, aes(x=Economy..logged.GDP.per.Capita., y=Health..Life.Expectancy.)) +
  geom_point() + # Show dots
  labs(title= "GDP vs Health Expectancy", x= "Economy (GDP/Capita)", y= "Life Expectancy") +
  geom_text(
    label= data$Country,
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T) + stat_smooth(method = "lm", se = FALSE)

hist(GDPvsHealthExpectancy$residuals)
 
```
This linear regression was intended to show the relationship between GDP and health and a line that emphisizes the positive correlation of linear graph.
The Histogram of the Residual can be used to check whether the variance is normally distributed. A symmetric bell-shaped histogram which is evenly distributed around zero indicates that the normality assumption is likely to be true. But since the histogram has a slight negative skew there is likely to be other factors contributing to the varience in the data.


```{r}

df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')
df <- select(df, COUNTRY, CODE)
colnames(df)[1] <- 'Country'
happiness_map_2021 <- merge(x = wh_2021_df, y = df, by = 'Country', all = TRUE)
happiness_map_2021 <- select(happiness_map_2021, Country, CODE, Happiness.Score)
happiness_map_2021 <- na.omit(happiness_map_2021)
  

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)
 
fig <- plot_geo(happiness_map_2021)
fig <- fig %>% add_trace(
  z = ~Happiness.Score, color = ~Happiness.Score, colors = 'Greens',
  text = ~Country, locations = ~CODE, marker = list(line = l)
)
fig <- fig %>% colorbar(title = 'Happiness Score', tickprefix = ':) ')
fig <- fig %>% layout(
  title = 'World Happiness 2021',
  geo = g
)

fig

```

This visualization is intended to show how a color gradient of the world happiness scale in the year 2021. This can be used to visually capture the major differences of the global west vs the global east through a color palette. 
